---
title: "Analyzing multiple MaxQuant files with VolcanoPlotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{multiple}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

To combine more than one MaxQuant file, you can use the `load_multiple_maxquant()` function to load them all into R.

The best approach is to place each `proteinGroups.txt` into its own subdirectory in the `Data` directory; the command will find and load any `proteinGroups.txt` in the `Data` directory.

The output of `load_multiple_maxquant()` is a list of data frames containing the combined data from all the files.

## Processing the data

### Specifiying groups

Using `process_multiple_maxquant()` with the list of data frames, you will be given the option of specifying which group(s) corresponds to group1 (right side of the volcano plot) and which group(s) corresponds to group2 (left side of the volcano plot). The function will then combine the data from all the files and return a single data frame that can be used for plotting.

In a simple example, two MaxQuant files have two groups each, "WT" and "Control". The runs in the both files are called WT.1, WT.2, WT.3, Control.1, Control.2, Control.3. The function presents the user with a list of the groups found in the files and asks which group(s) should be used for group1 and which group(s) should be used for group2. If we specify WT as group1 and Control as group2, the function will combine them. In doing so it will rename the runs to test.1, test.2, test.3, test.4, test5, test.6, control.1, control.2, control.3, control.4, control.5, control.6 to avoid any duplicate names.

In a more complicated example, in another set, the groups are called "WT" and "Control" in one file and "wildtype" and "control" in the other. The user can select from a list to specify the groups in group1 and those in group2. They would see a list of 1 WT, 2 Control, 3 wildtype, 4 control. If they select "1 3" for group1 and "2 4" for group2, the function will combine them and rename the runs to test.1, test.2, test.3, test.4, test5, test.6, control.1, control.2, control.3, control.4, control.5, control.6.

Note that the group names need to be unique and logical, for example if WT and Control are mislabelled in one file, or if names are shared between those in group1 and group2 then they cannot be combined. The solution here is to rerun MaxQuant with appropriate labelling.

In both of these examples, the user can bypass the selection step by specifying the groups in the function call, that is using `load_multiple_maxquant(group1 = "WT", group2 = "Control")` for the first example and `load_multiple_maxquant(group1 = c("WT", "wildtype"), group2 = c("Control", "control"))` for the second example.

```{r, eval = FALSE}
library(VolcanoPlotR)
my_list <- load_multiple_maxquant()
df_subset <- process_multiple_maxquant(data = my_list)
# or to bypass the selection step, specify the groups in the function call
df_subset <- process_multiple_maxquant(data = my_list, group1 = c("WT", "wildtype"), group2 = c("Control", "control"))
```

There is an additional argument that can be used to use the average of the ratios rather than ratio the averages. This is useful if datasets have been combined but they have differing intensities. Using `ratio = TRUE` will compare test.1 with control.1, test.2 with control.2, etc. and then average the ratios. Using `ratio = FALSE` will average the test samples and average the control samples and then compare the averages (this is the only option in the single file case). The default is `ratio = FALSE`.

## Plotting

The next steps are the same as if a single MaxQuant file had been used. Use `volcano_plot_maxquant()` or `pca_plot_maxquant()` or `mean_plot_maximum()` to generate the plots. `process_multiple_maxquant()` can be used as input to either of these functions.
