---
title: "Outputs from VolcanoPlotR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Outputs from VolcanoPlotR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

The main output is obviously, the volcano plot itself, which is returned as a ggplot object from the `volcano_plot_maxquant()` function. This can be customised further using ggplot2 functions, but there are other outputs that can be generated from the MaxQuant data.

# Text output

To generate a text output of the enriched proteins, you can set the `text_output` parameter to `TRUE` in the `volcano_plot_maxquant()` function. This will generate a text file containing the enriched proteins and their associated statistics. The output is saved as a tab-separated text file in the `text_output_dir` directory (default is "Output/Data/").

It contains a ranked list of proteins where the ranking is calculated by the manhattan distance from the origin (0,0) in the volcano plot. This list can be useful for other applications or for generating a table for publication.

# Alternative processing and statistical tests

## Changing the processing of MaxQuant data

By default, `process_maxquant()` will analysis LFQ intensity values. To use a different measurement, the `meas` parameter can be set to one of the other measurements in the MaxQuant output (e.g. "iBAQ", "Intensity", "MS/MS.counts").

The `baseval` parameter can be set to a value other than 0 to use a different base value for log2 transformation. The `width` and `downshift` parameters can be set to different values to change the imputation of missing values. The `seed` parameter can be set to a different value to change the random seed used for imputation.

## Statistical tests

The standard method for calculating p-values is via an unpaired two sample t-test assuming equal variance. To change to paired t-test, you can set the `paired` parameter to `TRUE` in the `process_maxquant()` function, similarly the `var.equal` parameter can be set to `FALSE` to use Welch's t-test instead of Student's t-test. The default settings mirror Perseus processing.

# Other plots

## Mean vs Mean plot

The function `mean_plot_maxquant()` can be used to generate a mean vs mean plot of the two groups being compared. This is useful for visualising the distribution of the data.

The plot is generated with similar parameters to the volcano plot. Some examples are shown below:

```{r mean_plot}
library(VolcanoPlotR)
# get the path to the proteinGroups.txt file included in the package
filepath <- system.file("extdata", "proteinGroups.txt", package = "VolcanoPlotR")
# get the filename from the path
filename <- basename(filepath)
# get the directory name
filedir <- dirname(filepath)
df <- load_maxquant(file = filename, datadir = filedir)
df <- process_maxquant(df, group1 = "WT", group2 = "Control")
# now we can generate the mean vs mean plot
mean_plot_maxquant(df)
# we can add a line to show the diagonal and start to label points
mean_plot_maxquant(df, xy_line = TRUE, label_points = "top_20")
# the parameters are the same as for the volcano plot, so we can change the colours and labels
mean_plot_maxquant(df, label_points = "5_10",
                   x_label = "Control (LFQ Log2)",
                   y_label = "Important group (LFQ Log2)",
                   vp_colours = c("0" = "#a0a000", "1" = "#808080",
                                  "2" = "#606060", "3" = "#00dd80",
                                  "4" = "#606060", "5" = "#800080"))
```

